---
title: "class12: Bioinformatics in drug discovery and design"
author: "Anusorn Mudla"
date: "5/10/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(bio3d)
# download the pdb file for 1hsg protein
file <- get.pdb("1hsg")
# read the pdb file
hiv <- read.pdb(file)
#print summary of the file
hiv
write.pdb(hiv,file = "1hsg.pdb")
```
Q1: What is the name of the two non protein resid values in this structure? What does resid correspond to and how would you get a listing of all reside values in this structure?
```{r}
#hiv$atom[!resid==PRO]
```

```{r}
protein <- atom.select(hiv,"protein",value = TRUE)
protein
write.pdb(protein,file = "1hsg_protein.pdb")
```
```{r}
ligand <- atom.select(hiv,"ligand",value = TRUE)
ligand
write.pdb(ligand,file = "1hsg_ligand.pdb")
```
*Add the hydrogen pdb files: both proetein and ligand and save them as pdbqt files

*generate the config.txt file

*perform the docking with AutoDock Vina using the terminal command

```{r}
res <- read.pdb("all.pdbqt", multi=TRUE)
res
write.pdb(res, "results.pdb")
```

compare our docking poses to the MERK drug structure
```{r}
ori <- read.pdb("1hsg_ligand.pdbqt")
rmsd(ori,res)
```

##Section 3: Exploring the conformational dynamics of proteins
```{r}
library(bio3d)
pdb <- read.pdb("1HEL")
modes <- nma(pdb) 
plot(modes, sse=pdb)
 # Visualize NMA results
# this will show the movement of the molecules
mktrj(modes, mode=7, file="nma_7.pdb")

```


