---
title: "Homework_class06"
author: "Anusorn Mudla"
date: "4/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Function Documentation
Description:

The function takes PDB protein ID code as a character type for the input and plot the atomic mobility of the alpha carbon in the A chain. The function allows users to compare how mobility in changes to the protein upon any treatment such as inhibitors. 

Usage:

plotpdb.mobil(protein_ID)

protein_ID is a string of four characters representing a PDB protein code.

```{r}
#install.packages("bio3d")
```

```{r}
library(bio3d)
```
## Function
```{r}
plotpdb.mobil <- function(protein_ID) {
# read the PDB data for the given protein
  full_data <- read.pdb(protein_ID)
# create a smaller PDB object containing only A chain and alpha carbon
  trim_data <- trim.pdb(full_data, chain="A", elety="CA")
# select the temperature factor "b" which indicates the mobility of the atom
  trim_data.b <- trim_data$atom$b
# plot the secondary structure of the protein
  plotb3(trim_data.b, sse=trim_data, typ="l", ylab="Bfactor")
}
```

## Test the function 
protein ID = "4AKE" and "1AKE"
```{r}
plotpdb.mobil("4AKE")
plotpdb.mobil("1AKE")

```

